<!-- question-type: inclass -->
### Exercise 3: Opening the auction database

So far, we have worked with an auction-level CSV file and bid-level JSON Lines data. Those files are useful, but they each give us only one view of the platform.

The database gives us several linked tables: bids, auctions, items, users, and orders. This lets us ask relational questions, such as which bidder appears in which seller's auctions, without loading every table into memory at once. In this exercise, the goal is to orient yourself to the database and create a bidder-level activity summary that we will reuse in the next exercises.

::: {.callout-warning collapse=true}
## Quarto gotcha: close your DuckDB connection before you Render {.unnumbered}

Quarto renders your document in a fresh R session. If you opened a DuckDB connection in the Console (or earlier chunk) and leave it open, that session may lock the `.duckdb` file, causing errors like "database is locked" or "unable to open database file."

You can fix this by following one of these routes:

* Manage the connection inside the document and close it at the end

```{r}
#| eval: false
library(DBI)
library(duckdb)
con <- dbConnect(duckdb(), dbdir = "data/auctions.duckdb")

# ... your analysis ...

dbDisconnect(con, shutdown = TRUE)
```

If you connected in the Console, run:

```{r}
#| eval: false
dbDisconnect(con, shutdown = TRUE)
```

before clicking Render.

(If you still see a lock, especially on Windows, restart R and render again.)
:::

**(a)**. Explain how a database differs from the CSV and JSON data sources we have used so far. Why is a database useful for this part of the shill-bidding investigation?

**(b)**. Connect to the database stored in `data/auctions.duckdb` and list the tables present in it by completing the code below.

```{r}
#| eval: false
con <- dbConnect(duckdb(), dbdir = "YOUR_FILENAME")

dbListTables(YOUR_CODE)
```

**(c)**. Inspect the tables. For each table, write a short note describing what one row appears to represent and which key column could link it to another table.

```{r}
#| eval: false
tbl(con, "bids")       |> glimpse()
tbl(con, "YOUR_TABLE") |> glimpse()
tbl(con, "YOUR_TABLE") |> YOUR_FUNCTION()
YOUR_FUNCTION(con, "YOUR_TABLE") |> YOUR_FUNCTION()
YOUR_FUNCTION(YOUR_OBJECT, "YOUR_TABLE") |> YOUR_FUNCTION()
```

**(d)**. The first review question is simple: which bidders are most active? Complete the code below to count each bidder's total number of bids and the number of auctions they have participated in.

```{r}
#| eval: false
bidder_activity <- 
    tbl(con, "bids") |>
    group_by(YOUR_VARIABLE) |>
    YOUR_VARIABLE(
        total_bids = YOUR_FUNCTION(),
        auctions_participated = YOUR_FUNCTION(YOUR_VARIABLE),
        .groups = "drop"
    ) |>
    collect()
```

**(e)**. Inspect `bidder_activity`. Why is high activity useful as a starting point for review, and why is it not enough to suggest shill bidding on its own?

<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

**Learning objective.** Students should understand why the investigation moves from files to a database, identify the main tables and keys, and create a bidder-level activity summary using `dplyr` on a DuckDB backend.

**Core concepts to highlight**

- A database holds multiple related tables rather than one flat file.
- `tbl(con, "...")` creates a lazy table; the database does the work until `collect()` brings results into R.
- Keys link the analysis path: bids use `auction_id`, auctions link to `item_id`, items identify `seller_id`, and orders identify `winner_id`.
- Activity is a useful triage signal, but active bidders can be legitimate collectors, resellers, or regular users.

**Suggested prompts**

- "Which table is at the bid-event level? Which table is at the auction-outcome level?"
- "Why do we use `n_distinct(auction_id)` rather than `n()` for auctions participated?"
- "What would you want to know before treating an active bidder as suspicious?"

**Common misunderstandings**

- Treating all tables as if they have the same unit of observation.
- Forgetting that database-backed `dplyr` code is translated to SQL.
- Assuming high activity is suspicious without considering legitimate repeat participation.

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a)**.

A database holds multiple related tables linked by keys such as `auction_id` and `item_id`; CSV and JSON files are usually single data sources. The database is useful here because shill-bidding questions often require joins: bids need to be linked to auctions, auctions to items, items to sellers, and orders to winners.

**(b)**.

```{r}
#| eval: false
con <- dbConnect(duckdb(), dbdir = "data/auctions.duckdb")

dbListTables(con)
```

**(c)**.

```{r}
#| eval: false
tbl(con, "bids")     |> glimpse()
tbl(con, "auctions") |> glimpse()
tbl(con, "users")    |> glimpse()
tbl(con, "items")    |> glimpse()
tbl(con, "orders")   |> glimpse()
```

Likely units of observation:

- `bids`: one bid event in one auction by one user.
- `auctions`: one auction and its linked item.
- `items`: one item, including category, reserve price, and seller.
- `orders`: one completed auction outcome and winner.
- `users`: one platform user.

**(d)**.

```{r}
#| eval: false
bidder_activity <- 
    tbl(con, "bids") |>
    group_by(user_id) |>
    summarise(
        total_bids = n(),
        auctions_participated = n_distinct(auction_id),
        .groups = "drop"
    ) |>
    collect()
```

**(e)**.

High activity helps identify bidders worth reviewing because coordinated accounts often need to bid repeatedly to influence prices. But high activity alone is not enough: a bidder may simply be a frequent buyer, a specialist in a category, or a reseller. We need relationship and outcome evidence before interpreting the behaviour as suspicious.

:::

:::

:::
<!-- END PROFILE:r-solutions -->
