<!-- question-type: prepare -->
### Exercise 1: Auction results baseline.

**(a).** Read the CSV file located at `data/auction_results.csv` and examine its structure by completing the code below:

```{r}
#| eval: false
# Load the data
auction_results <-
    YOUR_CODE("data/auction_results.csv")

# Examine the structure
YOUR_CODE(auction_results)
```

**(b)**. What does one row represent in the dataset? Which columns look most informative for understanding auction outcomes?

**(c)**. Produce 3 plots using the starter code below:

1. A histogram of `final_price`
2. A histogram of `num_bids`
3. A scatterplot of the relationship between `final_price` and `num_bids`.

```{r}
#| eval: false

p1 <-
    auction_results |>
    ggplot() +
    geom_histogram(aes(x = YOUR_CODE)) +
    theme_minimal()

p2 <-
    auction_results |>
    ggplot() +
    YOUR_CODE(aes(x = YOUR_CODE), n_bins = YOUR_CODE) +
    theme_minimal()

p3 <-
    auction_results |>
    ggplot() +
    YOUR_CODE(aes(x = YOUR_CODE,
                   y = YOUR_CODE)
               ) +
    geom_smooth(method = "lm") +
    theme_minimal()

p1 / p2 / p3
```




**(d)**. Provide a short explanation of your findings from (c). Are there any patterns that are suggestive of shill bidding?

**(e)**. Next, let's explore how certain auction outcomes vary across sellers. Complete the starter code below to produce summary statistics by seller:

```{r}
#| eval: false
seller_profile <- 
    auction_results |>
    group_by(seller_id) |>
    summarise(
        n_auctions    = YOUR_CODE(),
        median_price  = YOUR_CODE(final_price, na.rm = TRUE),
        median_bids   = YOUR_CODE(num_bids, na.rm = TRUE),
        share_missing_winner = YOUR_CODE(is.na(YOUR_CODE)),
        .groups = "drop"
    ) |>
    arrange(YOUR_CODE(n_auctions))
```

**(f)**. Based on the seller profiles in (e), are there any patterns suggestive of shill bidding? What further information would you need to confirm any suspicion?

<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective** 
Students should:

- Load and inspect an auctions-level CSV and correctly state the unit of observation.
- Create and interpret basic plots (histograms, scatter) linking `final_price` and `num_bids`.
- Distinguish suggestive patterns from evidence, and articulate what additional data would be needed to assess shill bidding.

✅   **Core Concepts to Highlight**

* Granularity: One row = one completed auction outcome.
* Right-skew: Expect skew in final_price and num_bids; medians are often more informative than means.
* Association ≠ proof: More bids often correlate with higher prices, but counterexamples are common.
* "Suggestive" vs "confirmatory": Patterns like high num_bids with low final_price, or high final_price with few bids, are flags—not verdicts.
* Seller aggregation: Summaries (n_auctions, median_price, median_bids, share_missing_winner) surface differences.

💬 **Suggested In-Class Prompts** (if needed)

What does one row represent? Say it in one sentence."

"Before plotting: predict the shape of the final_price histogram. Were you right?"

"Point to one dot in the scatter that doesn't fit the general trend. Give one benign and one suspicious explanation."

"winner_id is missing—name one data reason and one market reason."

"If you could request two extra fields to investigate shilling, what would they be and why?"

📌 **Common Misunderstandings**

"More bids must mean higher price." → Often, but not always; commodities, ceilings, or reserve mechanics can flatten prices.

Treating outliers as proof of shilling. → They are review candidates only.

Using means on skewed data. → Prefer medians/quantiles; acknowledge skew explicitly in commentary.

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a)**.

```{r}
# Load the data
auction_results <-
    read_csv("data/auction_results.csv")

# Examine the structure
glimpse(auction_results)
```

**(b)**.

One row = one auction outcome for a specific `item_id` run by a `seller_id`.

Most informative for outcomes:
`final_price` (price result), `num_bids` (competitive intensity) and `seller_id` (lets us compare performance across sellers).


**(c)**. 

```{r}
p1 <-
    auction_results |>
    ggplot() +
    geom_histogram(aes(x = final_price)) +
    theme_minimal()

p2 <-
    auction_results |>
    ggplot() +
    geom_histogram(aes(x = num_bids), bins = 20) +
    theme_minimal()

p3 <-
    auction_results |>
    ggplot(aes(
        x = num_bids,
        y = final_price)
        ) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal()

p1 / p2 / p3

```

**(d)**.

**(e)**.

```{r}
#| eval: false
seller_profile <- 
    auction_results |>
    group_by(seller_id) |>
    summarise(
        n_auctions    = n(),
        median_price  = median(final_price, na.rm = TRUE),
        median_bids   = median(num_bids, na.rm = TRUE),
        share_missing_winner = sum(is.na(winner_id)),
        .groups = "drop"
    ) |>
    arrange(desc(n_auctions))
```

**(f)**. 
Interpreting seller profiles & what else you'd need

Possibly suggestive seller patterns (from `n_auctions`, `median_price`, `median_bids`):

- High `median_bids` but low median_price across many auctions.
- Extreme `median_price` (much higher/lower than peers) given similar `median_bids`.
- Further information required to assess/confirm, such as:
    - Bid-level timelines (who bid, when, increments, rapid back-and-forth patterns).
    - Repeated bidder–seller ties (same bidder frequently active in the same seller's auctions).
    - Device/IP or account linkage between seller and active bidders (from JSON/telemetry).



:::

:::

:::
<!-- END PROFILE:r-solutions -->
